The galactic chemical evolution of carbon: Implications for stellar nucleosynthesis

### 2025-11-25 carbon_paper/tex/main.tex
carbon paper arxiv version

%%%%%%%%%%% JWJ' editing macros %%%%%%%%%%%
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\strike}[1]{{\color{red} \sout{#1}}}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{%
Daniel A. Boyea,$^{1, 2}$\thanks{%
Contact e-mail:~\href{mailto:danielaboyea@gmail.com}{danielaboyea@gmail.com}}
James W. Johnson,$^{2, 3, 4}$
and
David H. Weinberg$^{2,4}$
\\
% List of institutions
$^{1}$Department of Physics \& Astronomy, University of Victoria, 3800 Finnerty Road, Victoria, BC, V8P 5C2, Canada
\\
$^{2}$Department of Astronomy, The Ohio State University, 140 W. 18th Ave., Columbus, OH, 43210, USA
\\
$^{3}$The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA, 91101, USA
\\
$^{4}$Center for Cosmology \& Astroparticle Physics (CCAPP), The Ohio State University, 191 W. Woodruff Ave., Columbus, OH, 43210, USA
}

@@

% Abstract of the paper
\begin{abstract}
Carbon (C) is thought to be produced by both core collapse supernovae (CCSN) and asymptotic giant branch (AGB) stars, but the relative contributions of these two sources are uncertain. 
% goals
We investigate the astrophysical origin of C using models of Galactic chemical evolution (GCE) appropriate for the Milky Way disk.
We benchmark our results against APOGEE subgiant abundances.
% results
The trend between [C/Mg] and [Mg/H] is set by the total C yield as a function of metallicity. Observations indicate a gently rising [C/Mg] with [Mg/H], but AGB C production is predicted to decline with metallicity. 
Our sample therefore favours a scenario in which CCSN yields rise with metallicity to offset declining AGB C yields and drive a subtle increase in [C/Mg] with [Mg/H].
This result is consistent with massive star nucleosynthesis models incorporating rotation.
%
The [C/Mg]-[Mg/Fe] trend is sensitive to delayed enrichment and therefore constrains the amount of AGB C production.
Given the slope of this relation, we find that AGB stars likely account for 10--40 per cent of C at solar metallicity.
% nuance
Artificially shifting the AGB C yields to lower masses also improves agreement with the [C/Mg]-[Mg/Fe] trend, possibly indicating a discrepancy with stellar evolution predictions or our assumed Fe production rate.
\end{abstract}

@@

Predictions of C yields from stellar evolution and supernova (SN) models are uncertain for numerous reasons (see, e.g., the reviews by \citealt{romano+10, KL14}).
Some prominent unknowns include mass loss rates \citep{sukhbold+16, beasor+2020}, rotational mixing \citep{frischknecht+16, LC18}, nuclear reaction rates \citep{herwig+05, herwig+austin2004}, and convection \citep{chieffi2001, ventura+13, issa+25}.
Recent work additionally suggests that small changes in a star's initial mass of 
${\sim}0.01\,\Mo$  have significant effects on the evolution and explosion outcomes of massive stars \citep{bruenn+2023, vartanyan_burrows2023}. Most yield tables have far courser initial-mass resolution.  Furthermore, many stars are found in binary systems \citep{sana+12}, the effects of which are almost entirely
unexplored in stellar nucleosynthesis models \citep{farmer+21}.
Robust predictions of metal yields from stellar evolution models therefore remain elusive.

@@

While any abundance trend likely suffers from systematic biases, we find that these biases do not impact our qualitative conclusions. 
We have also consider subgiants and dwarfs with similar selection criteria from GALAH DR4 \citep{galah_dr4}, Gaia-ESO DR5.1 \citep{gaiaeso_dr5.1}, LAMOST \citep{lamost}, and mixing-corrected APOGEE giants \citep{vincenzo+21}. 
All of these samples show similar trends in \caah\ and \caafe{}, but with global systematic offsets of ${\sim} 0.1\,{\rm dex}$ and slight differences in the slope of each trend. Global offsets would affect our inferred total C trend with metallicity. However, given that the \caafe{} trend is similar among different surveys, we would recover similar results for the inferred relative contribution of AGB stars to C production (Section~\ref{sec:mcmc_results}). Our results should therefore be robust against systematic differences between these surveys.

@@

    \caption[]{
        \textbf{Left:} The integrated C yield from AGB stars, $\Ycagb$, as a function of metallicity for each yield table,  calculated $10\,{\rm Gyr}$ after a stellar population forms. (Note that the shape of the integrated yields arises from our linear interpolation of yields in $Z$. Changing to interpolation linear in $\log Z$ does not substantially affect our results.)
        \textbf{Right:} Cumulative C production as a function of age for a single stellar population of metallicity $\log Z/\Zo=-0.1$. The dashed black line shows the cumulative return fraction of type Ia supernovae ($\propto t^{-1.1}$) for comparison. The cumulative production of \aton{} reaches a minimum of -3 at a time of 0.3 Gyr.
}

@@

The IMF-averaged yields computed from the literature tables differ in normalization and metallicity dependence.  
The left panel of Fig.~\ref{fig:agb-ssp} shows IMF-averaged C yields for each AGB model as a function of metallicity, computed at time $\tau = 10$ Gyr after the formation of a single stellar population.
The normalizations span a factor of \about{2} (e.g., between $6\times 10^{-4}$ and $12 \times 10^{-4}$ at $Z=0.33\,\Zo$).
Additionally, the slope varies by a factor of ${\sim}3$.
Both \nugrid{} and \aton{} predict non-monotonic C yields with $Z$, but modelling systematics may exceed the scale of non-monotonic features.
These variations are due to different choices of reaction rates, convection treatments, and mass-loss rates (see discussion below and \citealt{james+23}).

@@

For our comparisons later, we define the AGB yield fraction, 
\begin{equation}\label{eq:f_agb}
    \fagb := \frac{\Ycagb(\Zo)}{\Ycc(\Zo) + \Ycagb(\Zo)}.
    \end{equation}
This quantity should approximately represent the fraction of the Sun's birth C abundance originating from AGB stars. Although, since the Sun was not enriched entirely by solar metallicity predecessors, the equivalence of $\fagb{}$ to the yield ratio is not exact.

\begin{table*}
	\centering
    \caption[]{For each AGB yield set, the IMF-averaged AGB C yield at solar metallicity $y_{\rm C, 0}^{\rm AGB}$, the fraction of solar C produced in the model $f_\odot^{\rm AGB}$, and the masses and metallicities each yield table is sampled on.
    $f_{\sun}^{\rm AGB}$ is calculated based on $y_{\rm C, \sun}^{\rm AGB}$ assuming the fiducial total C yield of $\Yct = 26.7\times 10^{-4}$.
    }

@@

Massive stars form $^{12}$C in their cores through the triple-$\alpha$ reaction (He-burning), which is later released to the ISM through winds and supernovae. Both material ejected through stellar winds and through CCSN contribute to C production, although we refer to CCSN and massive star enrichment interchangeably.

@@

CCSN models predict a wide range of C yields, spanning nearly a factor of ten, even at fixed metallicity.
Fig.~\ref{fig:y_cc} shows the integrated yields of massive star models from the literature. 
 The \cite{LC18} models, which include rotation, show that the induced mixing (e.g., \citealt{frischknecht+16}) can dramatically increase the magnitude and metallicity dependence of $\Ycc$. As we will later empirically show, CCSN C production needs to be strongly metallicity-dependent at $Z \approx \Zo$, which is consistent with \citepos{LC18} rapidly rotating models and to a lesser extent \citet{NKT13}.
Fig.~\ref{fig:y_cc} also shows the IMF-averaged AGB star yields from the \fruity\ model (in Fig.~\ref{fig:agb-ssp}). Especially at $Z\approx \Zo$, most CCSN models dominate AGB C production.

@@

We chose to adopt a linear yield model instead of a particular literature yield table, given the wide range of predictions for CCSN C. We parameterize the yield as
\begin{equation}\label{eq:y_cc}
    \Ycc =  \yo + \zcc \left(\frac{Z - \Zo}{\Zo}\right),
\end{equation}
where $\yo$ represents the yield at solar metallicity and $\zcc$ is the slope of the metallicity dependence. For our fiducial model, we adopt $\yo = 18.7\times 10^{-4}$ and $\zcc = 7.9 \times 10^{-4}$, leading to reasonable agreement with the \caah{}, \caafe{}, and gas-phase (Section~\ref{sec:gas}) relations. We present best-fitting models to subgiants only in Section~\ref{sec:mcmc_results}. We have also explored higher-order prescriptions (i.e., second-, third-, and fourth-order polynomials), finding improved agreement with the observed abundances with each successive free parameter.
The linear prescription likely does not fully capture C production by massive stars, but more complex models do not significantly alter our conclusions.

@@

We use the radial migration prescription described in appendix C of \citet{dubay+24} and summarized in Section 3 of \citet{james+24}. The total distance each star moves is sampled from a normal distribution with standard deviation $\sigma_R =  2.68\,{\rm kpc} \ (R/8\,{\rm kpc})^{0.61}\,(\tau / 8\,{\rm Gyr})^{0.33}$. Stars migrate from their initial to their final positions with a cube-root time-dependence. 
These parameters arise from \citepos{dubay+24} analytic approximation of the {\tt h227} simulation \citep{bird+21, christensen+12, zolotov12, loebman12, BZ14}.

Following \JJ{} (and more recently \citealt{james+24}), we control the slope of the radial metallicity gradient by adjusting the strength of the mass-loading in Galactic winds. The mass-loading, $\eta :=  \dot\Sigma_{\rm out} / \dot \Sigma_\star$, the ratio between the surface density rates of gas leaving the system and star formation, is set to
\begin{equation}\label{eq:mass_loading}
\eta(R) = \frac{y_{\rm Mg}^{\rm CC}}{Z_{\rm Mg}(R)} -1 + r %+ \tau_{\star} / \tau_{\rm sfh} 
\end{equation}
where 
\begin{equation}
    \log Z_{\rm Mg}(R) = \log Z_{\rm Mg,\ \sun} + 
    0.29 + 
    \begin{cases}
        -0.015(R-5) & R < 5 \\
        -0.09(R-5) & R \geq 5
    \end{cases}
\end{equation}
is adapted from \citet{hayden+14} assuming Mg is representative of $\alpha$ elements.
The recycling correction $r\approx 0.4$ accounts for the return of stellar envelopes back to the ISM (for a \citealt{kroupa01} IMF after 10 Gyr; see Section 2 from \citealt{WAF17}).
This choice of $\eta(R)$ results in a [$\alpha$/H] gradient consistent with APOGEE measurements \citep[e.g.,][]{hayden+14, weinberg+19, frinchaboy+13}.
Compared to \citet{james+21, james+24}, our metallicity gradient is flatter within 5 kpc (and our $\eta\to0$ at similar radius).
If we change our assumed yield scale, the values of $\eta$ will re-scale accordingly to maintain observationally-consistent chemical trends. Following \JJ, we neglect radial gas flows.

@@

Table~\ref{tab:fiducial_mod} summarizes our yield choices. 
Our solar abundance scale is from \citet{magg+22} with a 0.04 dex gravitational settling correction. Also from \citet{magg+22}, we adopt $\Zo = 0.0176$. We use the O, Mg, and Fe yields recommended by \citet{david_fe} based on measurements of the mean CCSN Fe yield \citep{rodriguez+21, rodriguez+23} and APOGEE chemical abundances.
Our AGB N yield is scaled from \citep[eq. 14]{james+23}, based on the differences in our assumed solar  N/O ratio. The CCSN N/O yield is additionally enhanced by a factor of 2.5 to account for the difference in median solar [N/Mg] values from \citet{vincenzo+21} and APOGEE subgiants. Following \citet{james+23}, we use a $t^{-1.1}$  delay time distribution (DTD) with a minimum delay-time of 150 Myr, based on comparisons of the SN Ia rate as a function of redshift and the cosmic SFR (e.g., \citealt{maoz+mannucci12}). We assume $2.2\times10^{-3}$ SN Ia events per unit mass of star formation with a mean $0.355\,\Mo$ of new Fe produced per supernova.
We use a \citet{kroupa01} initial mass function (IMF).

\begin{table}
	\centering
    \caption[]{Solar abundance scale and fiducial yields (in units of stellar population birth mass). See Section \ref{sec:agb} for the definition of \fruity. The solar abundance scale is \citet{magg+22} corrected upwards by 0.04 dex for gravitational settling and diffusion \citep{david_fe}.
    }
	\label{tab:fiducial_mod}

@@

Fig.~\ref{fig:c_evo} shows evolutionary tracks in \caah{} and \caafe{} in the fiducial model at 1\,kpc intervals of radius.
The \caah\ trend reaches an equilibrium state within \about{5}\,Gyr. \caafe\ tracks continue to evolve due to the long tail of the SN Ia DTD.

The evolution of chemical abundances can be understood as the interplay between different processes with different time-scales. 
 Initially, CCSN dominate enrichment. The [C/Mg] abundance ratio is set by the relative yield $\Ycc/\Ymg$. Due to the metallicity dependence of $\Ycc$, [C/Mg] increases with  [Mg/H]. [Mg/H], [Fe/H], and [C/H] all rise rapidly. [Mg/Fe] is initially high due to the delayed production time-scales of SN Ia Fe, as expected from the analytic models of \citet{WAF17}.
 Of order 100 Myr later, delayed sources enrich the ISM. AGB stars release C, raising [C/Mg], and SN Ia expel Fe, lowering [Mg/Fe]. 
 Several Gyr later, [Mg/H] reaches equilibrium. As the metal abundance stabilizes, so do metallicity-dependent CCSN yields, allowing C to reach equilibrium. [Fe/H] continues to rise slowly due to ongoing SN Ia from old stellar populations. 
 Finally, the system reaches approximate equilibrium and evolution slows. \

This evolution is predominantly driven by our chosen elemental yields, their
metallicity dependence, and their DTDs.
As we will demonstrate in Section~\ref{sec:agb_results} below, the final slope of the \caah\ trend depends mostly on the metallicity dependence of the total C yield. The \caafe{} trend instead reveals the nature of delayed C production.

@@

N production, also affected by the CNO cycle, is closely tied to C.
This paper builds upon \citet{james+23}, who used similar methods to constrain N yields. As an additional test of both yield models, we combine a rescaled version of their recommended N yields (see discussion in Section~\ref{sec:vice}) with our fiducial choice of C yields.
Fig.~\ref{fig:nitrogen} shows [C/N] versus [Mg/H] as predicted by our fiducial model. We are indeed able to closely match the observed [C/N]-[Mg/H] trend.

@@

The top row of Fig.~\ref{fig:zeta_f} shows models with varying strengths of the $\Ycc$ metallicity dependence, $\zcc$ (see Eq.~\ref{eq:y_cc}). With higher $\zcc$, the model predicts a steeper \caah\ trend, owing to the direct relationship between equilibrium [C/Mg] and yield ratios in Eq.~\ref{eq:c_mg_cc}. 
However, the \caafe{} trend is minimally affected by changes to $\zcc$ for stars at a given [Mg/H].
CCSN occur on much shorter time-scales than AGB and SN Ia enrichment.  The primary effect of increasing $\zcc$ in the \caafe{} plane, restricted to a metallicity slice, is to shift [C/Mg] uniformly.

The middle row of Fig.~\ref{fig:zeta_f} shows three models with different C AGB  fractions (Eq.~\ref{eq:f_agb}).
At fixed metallicity, the \caafe\ trend is principally sensitive to how much C and Fe come from delayed sources. In the limit $\fagb \to 0$, the [C/Mg] abundance will become nearly independent of [Mg/Fe]. As $\fagb$ increases, the steepness of the \caafe{} trend increases.

@@

All AGB yield tables predict a similarly shaped trend in the \caafe{} abundance space. The left panel of Fig.~\ref{fig:sims_degens} shows the best-fitting AGB yield models (see Section \ref{sec:mcmc} for methodology) for each yield table we consider. Although the individual DTDs of each AGB table appear substantially different (Fig.~\ref{fig:agb-ssp}), the predicted abundance trends do not show substantial differences. 
This outcome arises because each of our AGB models predict characteristic delay times for C production of order ${\sim}1$ Gyr, similar to the SN Ia production of Fe (see Figure 3), leading to a quasi-linear [C/Mg]-[Mg/Fe] relation in all cases.

However, all AGB tables predict a slightly concave down trend because C production by single stellar populations slows down significantly after a $\sim$few Gyr, while Fe production through SNe Ia continues up to ${\sim}10$ Gyr (see discussion in Section~\ref{sec:agb}). Our subgiant sample does not exhibit this level of concavity. These discrepancies are significant because Figure \ref{fig:sims_degens} plots the mean trend, which is precisely determined through the large sample size. One of the largest discrepancies between our models and data in the left panel is that \aton{} and \monash{} predict an almost flat \caafe{} trend for ${\rm [Mg/Fe]} \lesssim 0.1$.

>>>>

%%%%%%%%%%% JWJ' editing macros %%%%%%%%%%%
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\add}[1]{\red{#1}}
%\newcommand{\strike}[1]{\red{\sout{#1}}}
%\newcommand{\replace}[2]{\red{\sout{#1}} \red{#2}}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{%
Daniel A. Boyea,$^{1, 2, 3}$\thanks{%
Contact e-mail:~\href{mailto:danielaboyea@gmail.com}{danielaboyea@gmail.com}}
James W. Johnson,$^{4, 2, 3}$
and
David H. Weinberg$^{2,3}$
\\
% List of institutions
$^{1}$Department of Physics \& Astronomy, University of Victoria, 3800 Finnerty Road, Victoria, BC, V8P 5C2, Canada
\\
$^{2}$Department of Astronomy, The Ohio State University, 140 W. 18th Ave., Columbus, OH, 43210, USA
\\
$^{3}$Center for Cosmology \& Astroparticle Physics (CCAPP), The Ohio State University, 191 W. Woodruff Ave., Columbus, OH, 43210, USA\\
$^{4}$The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA, 91101, USA
}

@@

% Abstract of the paper
\begin{abstract}
Carbon (C) is thought to be produced by both core collapse supernovae (CCSN) and asymptotic giant branch (AGB) stars, but the relative contributions of these two sources are uncertain. 
% goals
We investigate the astrophysical origin of C using models of Galactic chemical evolution (GCE) appropriate for the Milky Way disk.
We benchmark our results against APOGEE subgiant abundances.
% results
The trend between [C/Mg] and [Mg/H] is set by the total C yield as a function of metallicity. Observations indicate a gently rising [C/Mg] with [Mg/H], but AGB C production is predicted to decline with metallicity. 
Our sample therefore favours a scenario in which CCSN yields rise with metallicity to offset declining AGB C yields and drive a subtle increase in [C/Mg] with [Mg/H].
This result is consistent with massive star nucleosynthesis models incorporating rotation.
%
The [C/Mg]-[Mg/Fe] trend is sensitive to delayed enrichment and therefore constrains the amount of AGB C production.
Given the slope of this relation, we find that AGB stars likely account for 10--40 per cent of C at solar metallicity.
% nuance
Artificially shifting the AGB C yields towards lower mass stars with longer lifetimes also improves agreement with the observed [C/Mg]-[Mg/Fe] trend, possibly indicating a discrepancy with stellar evolution predictions or our assumed Fe production rate.
\end{abstract}

@@

Predictions of C yields from stellar evolution and supernova (SN) models are uncertain for numerous reasons (see, e.g., the reviews by \citealt{romano+10, KL14}).
Some prominent unknowns include mass loss rates \citep{sukhbold+16, beasor+2020}, rotational mixing \citep{frischknecht+16, LC18}, nuclear reaction rates \citep{herwig+05, herwig+austin2004}, and convection \citep{chieffi2001, ventura+13, issa+25}.
Recent work additionally suggests that small changes in a star's initial mass of 
${\sim}0.01\,\Mo$  have significant effects on the evolution and explosion outcomes of massive stars \citep{bruenn+2023, vartanyan_burrows2023}. Most yield tables have far coarser initial-mass resolution.  Furthermore, many stars are found in binary systems \citep{sana+12}, the effects of which are almost entirely
unexplored in stellar nucleosynthesis models \citep{farmer+21}.
Robust predictions of metal yields from stellar evolution models therefore remain elusive.

@@

While any abundance trend likely suffers from systematic biases, we find that these biases do not impact our qualitative conclusions. 
We have also considered subgiants and dwarfs with similar selection criteria from GALAH DR4 \citep{galah_dr4}, Gaia-ESO DR5.1 \citep{gaiaeso_dr5.1}, LAMOST \citep{lamost}, and mixing-corrected APOGEE giants \citep{vincenzo+21}. 
All of these samples show similar trends in \caah\ and \caafe{}, but with global systematic offsets of ${\sim} 0.1\,{\rm dex}$ and slight differences in the slope of each trend. Global offsets would affect our inferred total C trend with metallicity. However, given that the \caafe{} trend is similar among different surveys, we would recover similar results for the inferred relative contribution of AGB stars to C production (Section~\ref{sec:mcmc_results}). Our results should therefore be robust against systematic differences between these surveys.

@@

    \caption[]{
        \textbf{Left:} The integrated C yield from AGB stars, $\Ycagb$, as a function of metallicity for each yield table,  calculated $10\,{\rm Gyr}$ after a stellar population forms. (Note that the shape of the integrated yields arises from our linear interpolation of yields in $Z$. Changing to linear interpolation in $\log Z$ does not substantially affect our results.)
        \textbf{Right:} Cumulative C production as a function of age for a single stellar population of metallicity $\log Z/\Zo=-0.1$. The dashed black line shows the cumulative return fraction of type Ia supernovae ($\propto t^{-1.1}$) for comparison. The cumulative production of \aton{} reaches a minimum of -3 at a time of 0.3 Gyr.
}

@@

The IMF-averaged yields computed from the literature tables differ in normalization and metallicity dependence.  
The left panel of Fig.~\ref{fig:agb-ssp} shows IMF-averaged C yields for each AGB model as a function of metallicity, computed at time $\tau = 10$ Gyr after the formation of a single stellar population.
The normalizations span a factor of \about{2} (e.g., between $6\times 10^{-4}$ and $12 \times 10^{-4}$ at $Z=0.33\,\Zo$).
Additionally, the slope varies by a factor of ${\sim}3$.
Both \nugrid{} and \aton{} predict slightly non-monotonic C yields with $Z$, but modelling systematics may exceed the scale of non-monotonic features.
These variations are due to different choices of reaction rates, convection treatments, and mass-loss rates (see discussion below and \citealt{james+23}).

@@

For our comparisons later, we define the AGB yield fraction, 
\begin{equation}\label{eq:f_agb}
    \fagb := \frac{\Ycagb(\Zo)}{\Ycc(\Zo) + \Ycagb(\Zo)}.
    \end{equation}
This quantity should approximately represent the fraction of the Sun's birth C abundance originating from AGB stars. However, since the Sun was not enriched entirely by solar metallicity predecessors, the equivalence of $\fagb{}$ to the yield ratio is not exact.

\begin{table*}
	\centering
    \caption[]{For each AGB yield set, the IMF-averaged AGB C yield at solar metallicity $y_{\rm C, 0}^{\rm AGB}$, the fraction of solar C produced in the model $f_\odot^{\rm AGB}$, and the masses and metallicities each yield table is sampled on.
    $f_{\sun}^{\rm AGB}$ is calculated based on $y_{\rm C, \sun}^{\rm AGB}$ assuming the fiducial total C yield of $\Yct = 26.7\times 10^{-4}$ and $\aagb=1$.
    }

@@

Massive stars form $^{12}$C in their cores through the triple-$\alpha$ reaction (He-burning), which is later released to the ISM through winds and supernovae. Both material ejected through stellar winds and through supernovae contribute to C production, but we generally use the term CCSN to refer to the sum of both modes of massive star enrichment.

@@

CCSN models predict a wide range of C yields, spanning nearly a factor of ten, even at fixed metallicity.
Fig.~\ref{fig:y_cc} shows the integrated yields of massive star models from the literature. At solar metallicity, \citepos{sukhbold+16} C yield is $\approx0.5$ dex higher than \cite[non-rotating]{LC18}, likely driven by much stronger (C-enhanced) mass loss in \citet{sukhbold+16} for stars more massive than $\gtrsim40\,\Mo$.
 The \cite{LC18} models, which include rotation, show that the induced mixing (e.g., \citealt{frischknecht+16}) can dramatically increase the magnitude and metallicity dependence of $\Ycc$. As we will later empirically show, CCSN C production needs to be strongly metallicity-dependent at $Z \approx \Zo$, which is consistent with \citepos{LC18} rapidly rotating models and to a lesser extent \citet{NKT13}.
Fig.~\ref{fig:y_cc} also shows the IMF-averaged AGB star yields from the \fruity\ model (in Fig.~\ref{fig:agb-ssp}). Especially at $Z\approx \Zo$, most CCSN models dominate AGB C production.

@@

We chose to adopt a linear yield model instead of a particular literature yield table, given the wide range of predictions for CCSN C. We parameterize the yield as
\begin{equation}\label{eq:y_cc}
    \Ycc =  \yo + \zcc \left(\frac{Z - \Zo}{\Zo}\right),
\end{equation}
where $\yo$ represents the yield at solar metallicity and $\zcc$ is the slope of the metallicity dependence. For our fiducial model, we adopt $\yo = 18.7\times 10^{-4}$ and $\zcc = 7.9 \times 10^{-4}$, leading to reasonable agreement with the \caah{}, \caafe{}, and gas-phase (Section~\ref{sec:gas}) relations. We present best-fitting models to subgiants only in Section~\ref{sec:mcmc_results}. We have also explored higher-order prescriptions (i.e., second-, third-, and fourth-order polynomials), finding improved agreement with the observed abundances with each successive free parameter.
The linear prescription likely does not fully capture C production by massive stars, but more complex models do not significantly alter our conclusions. 
As shown in Fig.~\ref{fig:y_cc}, the fiducial parameters imply a yield that agrees reasonably well with the model predictions at low $Z$ and a metallicity dependence in between that of \citeauthor{LC18}'s (\citeyear{LC18}, LC18) non-rotation and $v_{0}=300\,{\rm km\,s^{-1}}$ models.

@@

We use the radial migration prescription described in appendix C of \citet{dubay+24} and summarized in section 3 of \citet{james+24}. The total distance each star moves is sampled from a Gaussian distribution with standard deviation $\sigma_R =  2.68\,{\rm kpc} \ (R/8\,{\rm kpc})^{0.61}\,(\tau / 8\,{\rm Gyr})^{0.33}$. Stars migrate from their initial to their final positions with a cube-root time dependence. 
These parameters arise from \citepos{dubay+24} analytic approximation of the {\tt h227} simulation \citep{bird+21, christensen+12, zolotov12, loebman12, BZ14}.

Following \JJ{} (and more recently \citealt{james+24}), we control the slope of the radial metallicity gradient by adjusting the strength of the mass-loading in Galactic winds. The mass-loading, $\eta :=  \dot\Sigma_{\rm out} / \dot \Sigma_\star$, the ratio between the surface density rates of gas leaving the system and star formation, is set to
\begin{equation}\label{eq:mass_loading}
\eta(R) = \frac{y_{\rm Mg}^{\rm CC}}{Z_{\rm Mg}(R)} -1 + r %+ \tau_{\star} / \tau_{\rm sfh} 
\end{equation}
where 
\begin{equation}
    \log Z_{\rm Mg}(R) = \log Z_{\rm Mg,\ \sun} + 
    0.29 + 
    \begin{cases}
        -0.015(R-5) & R < 5 \\
        -0.09(R-5) & R \geq 5
    \end{cases}
\end{equation}
is adapted from \citet{hayden+14} assuming Mg is representative of $\alpha$ elements.
The recycling correction $r\approx 0.4$ accounts for the return of stellar envelopes back to the ISM (for a \citealt{kroupa01} IMF after 10 Gyr; see section 2 from \citealt{WAF17}).
This choice of $\eta(R)$ results in a [$\alpha$/H] gradient consistent with APOGEE measurements \citep[e.g.,][]{frinchaboy+13, hayden+14, weinberg+19}.
Compared to \citet{james+21, james+24}, our metallicity gradient is flatter within 5 kpc (and our $\eta\to0$ at similar radius).
If we change our assumed yield scale, the values of $\eta$ will re-scale accordingly to maintain observationally-consistent chemical trends. Following \JJ, we neglect radial gas flows.

@@

Table~\ref{tab:fiducial_mod} summarizes our yield choices. 
Our solar abundance scale is from \citet{magg+22} with a 0.04 dex gravitational settling correction. Also from \citet{magg+22}, we adopt $\Zo = 0.0176$. We use the O, Mg, and Fe yields recommended by \citet{david_fe} based on measurements of the mean CCSN Fe yield \citep{rodriguez+21, rodriguez+23} and APOGEE chemical abundances.
Our AGB N yield is scaled from \citep[eq. 14]{james+23}, based on the differences in the assumed solar N/O ratio and the median [N/Mg] or [N/O] value of each sample.  Following \citet{james+23}, we use a $t^{-1.1}$  delay time distribution (DTD) for SN Ia with a minimum delay-time of 150 Myr, based on comparisons of the SN Ia rate as a function of redshift and the cosmic SFR (e.g., \citealt{maoz+mannucci12}). We assume $2.2\times10^{-3}$ SN Ia events per unit mass of star formation with a mean $0.355\,\Mo$ of new Fe produced per supernova.
We use a \citet{kroupa01} initial mass function (IMF).

\begin{table}
	\centering
    \caption[]{Solar abundance scale and fiducial yields (in units of stellar population birth mass). See Section \ref{sec:agb} for the definition of \fruity. The solar abundance scale is \citet{magg+22} corrected upwards by 0.04 dex for gravitational settling and diffusion \citep{david_fe}. For $\Ycc$, the fiducial yield is given by Eq.~\ref{eq:y_cc} with $\yo=18.7\times10^{-4}$ and $\zcc=7.9\times10^{-4}$.
    }
	\label{tab:fiducial_mod}

@@

Fig.~\ref{fig:c_evo} shows evolutionary tracks in \caah{} and \caafe{} in the fiducial model at 1\,kpc intervals of radius (with C yields defined in Section~\ref{sec:nucleosynthesis}).
The \caah\ trend reaches an equilibrium state within \about{5}\,Gyr. \caafe\ tracks continue to evolve due to the long tail of the SN Ia DTD.

The evolution of chemical abundances can be understood as the interplay between different processes with different time-scales. 
 Initially, CCSN dominate enrichment. The [C/Mg] abundance ratio is set by the relative yield $\Ycc/\Ymg$. Due to the metallicity dependence of $\Ycc$, [C/Mg] increases with  [Mg/H]. [Mg/H], [Fe/H], and [C/H] all rise rapidly. [Mg/Fe] is initially high due to the delayed production time-scales of SN Ia Fe, as expected from the analytic models of \citet{WAF17}.
 Of order 100 Myr later, delayed sources enrich the ISM. AGB stars release C, raising [C/Mg], and SN Ia release Fe, lowering [Mg/Fe]. 
 Several Gyr later, [Mg/H] reaches equilibrium. As the metal abundance stabilizes, so do metallicity-dependent CCSN yields, allowing C to reach equilibrium. [Fe/H] continues to rise slowly due to ongoing SN Ia from old stellar populations. 
 Finally, the system reaches approximate equilibrium and evolution slows. \

This evolution is predominantly driven by our chosen elemental yields, their
metallicity dependence, and their DTDs.
As we will demonstrate in Section~\ref{sec:agb_results} below, the final locus of the \caah\ trend (the end points of the curves in the left panel) depends mostly on the metallicity dependence of the total C yield. The \caafe{} trend instead reveals the nature of delayed C production.

@@

N production, also affected by the CNO cycle, is closely tied to C.
This paper builds upon \citet{james+23}, who used similar methods to constrain N yields. As an additional test of both yield models, we combine a rescaled version of their recommended N yields (see discussion in Section~\ref{sec:vice}) with our fiducial choice of C yields. The rescaling is needed because we assume a lower $\Ymg$ and our sample has a higher median [N/Mg] than in \citet{james+23}.
Fig.~\ref{fig:nitrogen} shows [C/N] versus [Mg/H] as predicted by our fiducial model. We are indeed able to closely match the observed [C/N]-[Mg/H] trend.

@@

The top row of Fig.~\ref{fig:zeta_f} shows models with varying strengths of the $\Ycc$ metallicity dependence, $\zcc$ (see Eq.~\ref{eq:y_cc} and parameters in Table~\ref{tab:model_parameters}). With higher $\zcc$, the model predicts a steeper \caah\ trend, owing to the direct relationship between equilibrium [C/Mg] and yield ratios in Eq.~\ref{eq:c_mg_cc}. 
However, the \caafe{} trend is minimally affected by changes to $\zcc$ for stars at a given [Mg/H].
CCSN occur on much shorter time-scales than AGB and SN Ia enrichment.  The primary effect of increasing $\zcc$ in the \caafe{} plane, restricted to a metallicity slice, is to shift [C/Mg] uniformly.

The middle row of Fig.~\ref{fig:zeta_f} shows three models with different C AGB  fractions (Eq.~\ref{eq:f_agb}; parameters in Table~\ref{tab:model_parameters}).
At fixed metallicity, the \caafe\ trend is principally sensitive to how much C and Fe come from delayed sources. In the limit $\fagb \to 0$, the [C/Mg] abundance will become nearly independent of [Mg/Fe]. As $\fagb$ increases, the steepness of the \caafe{} trend increases.

@@

All AGB yield tables predict a similarly shaped trend in the \caafe{} abundance space. The left panel of Fig.~\ref{fig:sims_degens} shows the best-fitting yield combinations for each AGB table we consider (solving for $\aagb$, $\yo$, and $\zeta$ in Section \ref{sec:mcmc}; yield parameters in Table~\ref{tab:model_parameters}). Although the individual DTDs of each AGB table appear substantially different (Fig.~\ref{fig:agb-ssp}), the predicted abundance trends do not show substantial differences. 
This outcome arises because each of the AGB models predicts characteristic delay times for C production of order ${\sim}1$ Gyr, similar to the SN Ia production of Fe (see Fig.~\ref{fig:agb-ssp}), leading to a quasi-linear [C/Mg]-[Mg/Fe] relation in all cases.

However, all AGB tables predict a slightly concave down trend because C production by single stellar populations slows down significantly after a few Gyr, while Fe production through SNe Ia continues up to ${\sim}10$ Gyr (see discussion in Section~\ref{sec:agb}). Our subgiant sample does not exhibit this level of concavity. These discrepancies are significant because Fig.~\ref{fig:sims_degens} plots the mean trend, which is precisely determined through the large sample size. One of the largest discrepancies between our models and data in the left panel is that \aton{} and \monash{} predict an almost flat \caafe{} trend for ${\rm [Mg/Fe]} \lesssim 0.1$.
